### 2012-12-21 inst/examples/myers-exploration.Rmd
corrected iterations

```{r parallel}
require(snowfall)
sfInit(cpu=2)
sfExportAll()
sfLibrary(pdgControl)
sfLibrary(nonparametricbayes)
sfLibrary(reshape2)
sfLibrary(ggplot2)
sfLibrary(data.table)
sfLibrary(tgp)
sfLibrary(kernlab)
sfLibrary(MCMCpack)
sfLibrary(plyr)
```

```{r }
yields <- 
  sfLapply(1:20, function(j){

x_0_observed <- allee + x_grid[5]
seed <- round(runif(1) * 1e6)
set.seed(seed)
print(seed)

## @knitr sim-obs
Tobs <- 50
x <- numeric(Tobs)
x[1] <- x_0_observed
for(t in 1:(Tobs-1))
  x[t+1] = z_g() * f(x[t], h=0, p=p)
qplot(1:Tobs, x)

## @knitr lag-data
obs <- data.frame(x=c(0,x[1:(Tobs-1)]),y=c(0,x[2:Tobs]))

## @knitr par-est
estf <- function(p){
  mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
  -sum(dlnorm(obs$y, mu, p["s"]), log=TRUE)
}
o <- optim(par = c(r=1,K=mean(x),s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3))
f_alt <- Ricker
p_alt <- c(o$par['r'], o$par['K'])
sigma_g_alt <- o$par['s']

## @knitr gp-fit
gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
          m0r1=FALSE, corr="exp", trace=TRUE, 
          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")

## @knitr gp-data
V <- gp$ZZ.ks2
Ef = gp$ZZ.km
tgp_dat <- data.frame(x   = gp$XX[[1]], 
                      y   = gp$ZZ.km, 
                      ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
                      ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))

## @knitr gp-plot
true <- sapply(x_grid, f, 0, p)
est <- sapply(x_grid, f_alt, 0, p_alt)
models <- data.frame(x=x_grid, GP=tgp_dat$y, Parametric=est, True=true)
models <- melt(models, id="x")
names(models) <- c("x", "method", "value")
gp_plot <- ggplot(tgp_dat)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
  geom_point(data=obs, aes(x,y), alpha=0.8) + 
  xlab(expression(X[t])) + ylab(expression(X[t+1])) +
  scale_colour_manual(values=cbPalette)
print(gp_plot)

## @knitr gp-posteriors
hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
prior_curves <- ddply(posteriors, "variable", function(dd){
  grid <- seq(min(dd$value), max(dd$value), length = 100)
  data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
})
posteriors_plot <- ggplot(posteriors) + 
  geom_histogram(aes(x=value, y=..density..), lwd=2) +
  geom_line(data=prior_curves, aes(x=value, y=density), col="red", lwd=2) +
  facet_wrap(~ variable, scale="free")
#print(posteriors_plot)

## @knitr gp-opt
matrices_gp <- gp_transition_matrix(Ef, V, x_grid, h_grid)
opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)

## @knitr true-opt
matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)

## @knitr est-opt
matrices_estimated <- f_transition_matrix(f_alt, p_alt, x_grid, h_grid, sigma_g_alt)
opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)

## @knitr policy_plot
policies <- melt(data.frame(stock=x_grid, 
                            GP = x_grid[opt_gp$D[,1]], 
                            Parametric = x_grid[opt_estimated$D[,1]],
                            True = x_grid[opt_true$D[,1]]),
                 id="stock")
names(policies) <- c("stock", "method", "value")
policy_plot <- ggplot(policies, aes(stock, stock - value, color=method)) +
  geom_line(lwd=2, alpha=0.8) + 
  xlab("stock size") + ylab("escapement")  +
  scale_colour_manual(values=cbPalette)
print(policy_plot)

## @knitr stationary_policy_only
m <- sapply(1:OptTime, function(i) opt_gp$D[,1])
opt_gp$D <- m
mm <- sapply(1:OptTime, function(i) opt_true$D[,1])
opt_true$D <- mm
mmm <- sapply(1:OptTime, function(i) opt_estimated$D[,1])
opt_estimated$D <- mmm

## @knitr simulate
set.seed(1)
sim_gp <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit))
set.seed(1)
sim_true <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit))
set.seed(1)
sim_est <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_estimated$D, z_g, profit=profit))

## @knitr tidy
dat <- list(GP = sim_gp, Parametric = sim_est, True = sim_true)
dat <- melt(dat, id=names(dat[[1]][[1]]))
dt <- data.table(dat)
setnames(dt, c("L1", "L2"), c("method", "reps"))

## @knitr sim-fish
fish_plot <- ggplot(dt) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
  scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))+
  ggtitle(paste("rep", j, "seed", seed))
print(fish_plot)

## @knitr sim-harvest
harvest_plot <- ggplot(dt) +
  geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
  scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1))) +
  ggtitle(paste("rep", j, "seed", seed))
print(harvest_plot)

## @knitr costs
profits <- dt[, sum(profit), by = c("reps", "method")]
means <- profits[, mean(V1), by = method]
sds <- profits[, sd(V1), by = method]
yield <- cbind(means, sd = sds$V1)
yield

  yield
})
```

>>>>

```{r}

sim_obs <- function(Xo, z_g, f, p, Tobs = 40, seed = 1){
  x <- numeric(Tobs)
  x[1] <- Xo
  set.seed(seed)
  for(t in 1:(Tobs-1))
    x[t+1] = z_g() * f(x[t], h=0, p=p)
  data_plot <- qplot(1:Tobs, x) + ggtitle(paste("seed", seed))
  print(data_plot)
  obs <- data.frame(x=c(0,x[1:(Tobs-1)]),y=c(0,x[2:Tobs]))
}

par_est <- function(obs){
  estf <- function(p){
    mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
    llik <- -sum(dlnorm(obs$y, mu, p["s"]), log=TRUE)
    if(!is.numeric(llik) | is.nan(llik) | !(llik < Inf)){
      warning("possible error in llik")
      llik <- 1e30
    }
    llik
  }
  o <- optim(par = c(r=1,K=mean(obs$x),s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3))
  f_alt <- Ricker
  p_alt <- c(o$par['r'], o$par['K'])
  sigma_g_alt <- o$par['s']
  list(f_alt = f_alt, p_alt = p_alt, sigma_g_alt = sigma_g_alt)
}
optimal_policy <- function(gp, f, f_alt, p, p_alt, x_grid, h_grid, sigma_g, sigma_g_alt, delta, xT, profit, reward, OptTime){
  matrices_gp <- gp_transition_matrix(gp$ZZ.ks2, gp$ZZ.km, x_grid, h_grid)
  opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
  matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
  opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
  matrices_estimated <- f_transition_matrix(f_alt, p_alt, x_grid, h_grid, sigma_g_alt)
  opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
  list(gp_D = opt_gp$D[,1], true_D = opt_true$D[,1], est_D = opt_estimated$D[,1])
}

simulate_opt <- function(OPT, f, p, x_grid, h_grid, K, z_g, profit){
  gp_D <- sapply(1:OptTime, function(i) OPT$gp_D)
  true_D <- sapply(1:OptTime, function(i) OPT$true_D)
  est_D <- sapply(1:OptTime, function(i) OPT$est_D)
  set.seed(1)
  sim_gp <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, gp_D, z_g, profit=profit))
  set.seed(1)
  sim_true <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, true_D, z_g, profit=profit))
  set.seed(1)
  sim_est <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, est_D, z_g, profit=profit))
  dat <- list(GP = sim_gp, Parametric = sim_est, True = sim_true)
  dat <- melt(dat, id=names(dat[[1]][[1]]))
  dt <- data.table(dat)
  setnames(dt, c("L1", "L2"), c("method", "reps")) 
  dt
}

sim_plots <- function(dt, seed=1){
  fish_plot <- ggplot(dt) + 
    geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))+
    ggtitle(paste("seed", seed))
  print(fish_plot)
  harvest_plot <- ggplot(dt) +
    geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1))) +
    ggtitle(paste("seed", seed))
  print(harvest_plot)
}

profits_stats <- function(dt){
  profits <- dt[, sum(profit), by = c("reps", "method")]
  means <- profits[, mean(V1), by = method]
  sds <- profits[, sd(V1), by = method]
  yield <- cbind(means, sd = sds$V1)
  yield
}
```

```{r parallel, message=FALSE, warning=FALSE, results='hide', eval=FALSE, include=FALSE, echo=FALSE}
require(snowfall)
sfInit(par=TRUE, cpu=8)
sfExportAll()
sfLibrary(pdgControl)
sfLibrary(nonparametricbayes)
sfLibrary(reshape2)
sfLibrary(ggplot2)
sfLibrary(data.table)
sfLibrary(tgp)
sfLibrary(kernlab)
sfLibrary(MCMCpack)
sfLibrary(plyr)
```

```{r }
seed <- round(runif(2) * 1e6)
seed
yields <- lapply(seed, 
function(seed_i){
  Xo <- allee + x_grid[5]
  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
  est <- par_est(obs)
  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
          meanfn="constant", bprior="b0", BTE=c(2000,6000,2),
          m0r1=FALSE, corr="exp", trace=FALSE, 
          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
  OPT <- optimal_policy(gp, f, est$f_alt, p, est$p_alt, x_grid, h_grid, sigma_g, est$sigma_g_alt, delta, xT, profit, reward, OptTime)
  dt <- simulate_opt(OPT, f, p, x_grid, h_grid, K, z_g, profit)
  sim_plots(dt, seed=seed_i)
  profits_stats(dt)
})
```
